% !TEX root = ../thesis.tex

%TC:ignore

\begin{publications}
  \item {\bf{Li Y}}, Lu H. Natural Image Matting via Guided Contextual Attention. The AAAI Conference on Artificial Intelligence (AAAI). 2020. (CCF A)
  \item {\bf{Li Y}}, Zhang J, Zhao W, Weihao J, Lu H. Inductive Guided Filter: Real-time Deep Image Matting with Weakly Annotated Masks on Mobile Devices. IEEE International Conference on Multimedia and Expo (ICME). 2020  (CCF B)
  \item {\bf{Li Y}}, Lu H. On Multi-modal Fusion Learning in constraint propagation. Information Sciences. 2018 Sep 1;462:204-17. (CCF B)
  \item {\bf{Li Y}}, Chen J, Zhao Y, Lu H. Adaptive affinity matrix for unsupervised metric learning. IEEE International Conference on Multimedia and Expo (ICME). 2016 Jul 11 (pp. 1-6). IEEE. (CCF B)
  \item Zhang J, Huang Y, {\bf{Li Y}}, Zhao W, Zhang L. Multi-Attribute Transfer via Disentangled Representation. Proceedings of the AAAI Conference on Artificial Intelligence (AAAI). 2019 Jul 17 (Vol. 33, pp. 9195-9202). (CCF A)
  \item Zhang J, Niu L, Yang D, Kang L, {\bf{Li Y}}, Zhao W, Zhang L. GAIN: Gradient Augmented Inpainting Network for Irregular Holes. InProceedings of the 27th ACM International Conference on Multimedia (ACM MM). 2019 Oct 15 (pp. 1870-1878). (CCF A)
  \item Li Z, {\bf{Li Y}}, Lu H. Improve Image Captioning by Self-attention. InInternational Conference on Neural Information Processing (ICONIP). 2019 Dec 12 (pp. 91-98). Springer, Cham. (CCF C)
  \item Chen J, {\bf{Li Y}}, Lu H. Online self-organizing hashing. IEEE International Conference on Multimedia and Expo (ICME). 2016 Jul 11 (pp. 1-6). IEEE. (CCF B)
  \item Zhao Y, {\bf{Li Y}}, Shao Z, Lu H. Lsod: Local sparse orthogonal descriptor for image matching. Proceedings of the 24th ACM international conference on Multimedia (ACM MM). 2016 Oct 1 (pp. 232-236). (CCF A)
\end{publications}

\begin{publications*}
  \item 第一作者. Natural Image Matting via Guided Contextual Attention. The AAAI Conference on Artificial Intelligence (AAAI). 2020. (CCF A)
  \item 第一作者. Inductive Guided Filter: Real-time Deep Image Matting with Weakly Annotated Masks on Mobile Devices. IEEE International Conference on Multimedia and Expo (ICME). 2020. accepted. (CCF B)
  \item 第一作者. On Multi-modal Fusion Learning in constraint propagation. Information Sciences. 2018 Sep 1;462:204-17. (CCF B)
  \item 第一作者. Adaptive affinity matrix for unsupervised metric learning. IEEE International Conference on Multimedia and Expo (ICME). 2016 Jul 11 (pp. 1-6). IEEE. (CCF B)
  \item 第三作者. Multi-Attribute Transfer via Disentangled Representation. Proceedings of the AAAI Conference on Artificial Intelligence (AAAI). 2019 Jul 17 (Vol. 33, pp. 9195-9202). (CCF A)
  \item 第五作者. GAIN: Gradient Augmented Inpainting Network for Irregular Holes. InProceedings of the 27th ACM International Conference on Multimedia (ACM MM). 2019 Oct 15 (pp. 1870-1878). (CCF A)
  \item 第二作者. Improve Image Captioning by Self-attention. InInternational Conference on Neural Information Processing (ICONIP). 2019 Dec 12 (pp. 91-98). Springer, Cham. (CCF C)
  \item 第二作者. Online self-organizing hashing. IEEE International Conference on Multimedia and Expo (ICME). 2016 Jul 11 (pp. 1-6). IEEE. (CCF B)
  \item 第二作者. Lsod: Local sparse orthogonal descriptor for image matching. Proceedings of the 24th ACM international conference on Multimedia (ACM MM). 2016 Oct 1 (pp. 232-236). (CCF A)

\end{publications*}

%TC:endignore
